---
title: "gasToken analysis"
author: "Timothée Henry"
date: "November 5th, 2020"
output:
  html_document:
    self_contained: false
    lib_dir: libs
---

<style>
div.blue { background-color:#e6f0ff; border-radius: 3px; padding: 5px;}
</style>

# Data source  
    
gasToken
pair gasToken / WETH
https://info.uniswap.org/pair/0x27c64bdca05d79f6ee32c3e981dc5153d9d794cd

```{txt}
SELECT block_timestamp, reserve0, reserve1
FROM blockchain-etl.ethereum_uniswap.UniswapV2Pair_event_Sync
where contract_address="0x27c64bdca05d79f6ee32c3e981dc5153d9d794cd"
and date(block_timestamp)>="2020-01-01"
```


```{r}
require(data.table)
t=fread("/Users/henry/Desktop/perso/eth0/gasprice/gasToken-weth-pair-2.csv")
t[, reserve1 := as.numeric(reserve1)]
t[, priceGasToken := reserve1/reserve0 * 400 / 1e18 * 100]
t[, datetime := strptime(substr(block_timestamp, 1, 16), "%Y-%m-%d %H:%M")]
t = t[order(datetime)]
```


# Compare gasToken price with gasPrice
```{txt}
SELECT SUBSTR(STRING(block_timestamp), 1, 13) as date_time, avg(gas_price) as gas_price
FROM bigquery-public-data.crypto_ethereum.transactions
where date( block_timestamp)>="2020-05-18"
group by date_time
```

```{r, fig.width=14, fig.height=16}
g =fread("/Users/henry/Desktop/perso/eth0/gasprice/gasprice-june-nov.csv")
g = g[order(date_time)]
g[, datetime := strptime(date_time, "%Y-%m-%d %H")]
g[, gas_price := round(gas_price/1e9, 2)]
library(zoo)
# moving average over last 30 data points (1 hour or so)
g[, gas_price_ma := frollmean(gas_price, 30)]

# volume of trade, from swaps on same pair
swaps = fread("/Users/henry/Desktop/perso/eth0/gasprice/gasToken-weth-pair-swaps.csv")
swaps = swaps[order(block_timestamp)]
swaps = swaps[, list(volume = sum(as.numeric(amount1In)/1e18 * 400)), by=substr(block_timestamp, 1, 10)]

# ===================================
# PLOT
par(mfrow=c(3,1), mar = c(2, 2, 2, 2))

# First plot
g[, plot(datetime, gas_price_ma, ylim=c(0,300), col="blue", type="l", main="gasPrice vs gasToken_price")]
for(i in seq(0,300,10)) abline(h=i, col="lightgray")
g[, lines(datetime, gas_price_ma, ylim=c(0,300), col="blue")]
t[, lines(datetime, priceGasToken, col="red")]
legend("topleft", legend=c("gasPrice", "gasToken price on Uniswap GTK-WETH"),
       col=c("blue", "red"), lty=1:1, cex=1.4)

# Second plot
# show reserves on uni pool
t[, plot(datetime, reserve0/1000, type="l", col="orange", main="GTK-WETH Uniswap pair reserves")]
for(i in seq(0,200000,10000)) abline(h=i, col="lightgray")
t[, lines(datetime, reserve0/1000, col="orange")]
t[, lines(datetime, reserve1/1e18, col="green")]
legend("topleft", legend=c("GTK reserves / 1e3", "WETH reserves / 1e18"),
       col=c("orange", "green"), lty=1:1, cex=1.4)

# add volume of trades
swaps[, plot(as.Date(substr), volume, type="l", main="Volume of trade in USD on Uniswap GTK-WETH")]

```

# Comparing gasToken value and prices (on Uniswap, or minting)

gasToken value is given by the javascript code at https://gastoken.io/

MINT_BASE = 32254;
MINT_TOKEN = 36543;
FREE_BASE = 14154;
FREE_TOKEN = 6870;
REIMBURSE = 24000;

The value is the part REIMBURSE which is 24000 gas, which depends on the gas price at the time it is used.
The price of the gasToken depends if it is minted, or bought on Uniswap for example.

Let's calculate those, and plot:
```{r}
g[, x := substr(datetime, 1, 13)]
t[, x := substr(datetime, 1, 13)]

setkey(g, x)
setkey(t, x)

r = t[g]
r = r[!is.na(priceGasToken)][!is.na(gas_price_ma)]

r[, gasToken_Value := 24000 * gas_price_ma * 1e9 / 1e18 * 400]
r[, gasToken_Price_Mint := 36543 * gas_price_ma * 1e9 / 1e18 * 400]
r[, gasToken_Price_Uniswap := reserve1/reserve0 / 1e18 * 400]

par(mfrow=c(1,1), mar = c(2, 2, 2, 2))
r[, plot(datetime, gasToken_Value, col="green", type="l", main="gasToken value vs prices")]
r[, lines(datetime, gasToken_Price_Mint, col="red")]
r[, lines(datetime, gasToken_Price_Uniswap, col="orange")]
legend("topleft", legend=c("gasToken value", "gasToken price when minted", "gasToken price on Uniswap"),
       col=c("green", "red","orange"), lty=1:1, cex=0.8)


```


