---
title: "Uniswap UNI token transactions analysis"
author: "Timothée Henry"
date: "September 22nd, 2020"
output:
  html_document:
    self_contained: false
    lib_dir: libs
---

<style>
div.blue { background-color:#e6f0ff; border-radius: 3px; padding: 5px;}
</style>



# Objective


<div class = "blue">

The UNI token was introduced on 17th Septembre 2020
Refer to the Uniswap blog post for details:
https://uniswap.org/blog/uni/

The objective of this page is to analyze the **UNI token transactions**, and try to understand what is driving its price.

For this  we will try to answer the following questions:

* Question 1: who are the main holders?
* Question 2: what percentage of the airdrop was claimed so far?
* Question 3: did the receivers of airdrop keep their tokens?
* Question 4: which are the main exchanges?
* Question 5: who are the main buyers and sellers?
* Question 6: what is driving the price?

</div>

# Data source  

<div class = "blue">All data was downloaded using BigQuery</div>
The following queries were used:

To get token transfers on token UNI (0x1f9840a85d5af5bf1d1762f925bdaddc4201f984)
https://etherscan.io/address/0x1f9840a85d5af5bf1d1762f925bdaddc4201f984
```{txt}
SELECT * FROM bigquery-public-data.crypto_ethereum.token_transfers
WHERE token_address = "0x1f9840a85d5af5bf1d1762f925bdaddc4201f984"
```

and to get swaps on the ETH-UNI Uniswap exchange (contract 0xd3d2e2692501a5c9ca623199d38826e513033a17)
https://uniswap.info/pair/0xd3d2e2692501a5c9ca623199d38826e513033a17
https://etherscan.io/address/0xd3d2e2692501a5c9ca623199d38826e513033a17
```{txt}
SELECT * FROM blockchain-etl.ethereum_uniswap.UniswapV2Pair_event_Swap
where contract_address = "0xd3d2e2692501a5c9ca623199d38826e513033a17"
```


```{r}
require(data.table)
library(zoo)

t = fread("/Users/henry/Desktop/perso/eth/uniswap/uniswap/uni_transactions_22sep.csv")
s = fread("/Users/henry/Desktop/perso/eth/uniswap/uniswap/uni_swaps_22sep.csv")

# convert time
s[, date_time_str := substr(block_timestamp, 1, 16)]
#t[, date_time := strptime(date_time_str, "%Y-%m-%d %H:%M")]
s[, date_time := as.POSIXct(date_time_str, "%Y-%m-%d %H:%M", tz = "GMT")]
s = s[order(date_time)]
s[, amount0In := as.numeric(amount0In) / 1e18]
s[, amount1In := as.numeric(amount1In) / 1e18]
s[, amount0Out := as.numeric(amount0Out) / 1e18]
s[, amount1Out := as.numeric(amount1Out) / 1e18]
s = s[, list(date_time, amount0In, amount1In, amount0Out, amount1Out)]

# deduce UNI/ETH conversion rate
s[amount0In != 0, r1 :=  amount1Out / amount0In]
s[amount0In != 0, volume :=  amount1Out]

s[amount0Out != 0, r1 :=  amount1In / amount0Out]
s[amount0Out != 0, volume :=  amount1In]

#s[, r1_mavg := frollmean(r1, 300)]
s[, r1_mavg := frollsum(r1 * volume, 50) / frollsum(volume, 50)]

# simplify value as numeric in token
t[, value := as.numeric(value) / 1e18]

# keep only useful columns
t = t[, list(block_timestamp, value, from_address, to_address)]

# convert time
t[, date_time_str := substr(block_timestamp, 1, 16)]
#t[, date_time := strptime(date_time_str, "%Y-%m-%d %H:%M")]
t[, date_time := as.POSIXct(date_time_str, "%Y-%m-%d %H:%M", tz = "GMT")]

t = t[, list(date_time, value, from_address, to_address)]

t = t[order(date_time)]
```

# Question 1: who are the main holders?

This question is answered in the following etherscan page:
https://etherscan.io/token/0x1f9840a85d5af5bf1d1762f925bdaddc4201f984#balances

But we can check this by ourselves:

```{r}

t_from = t[, list(date_time, address = from_address, value = -value)]
t_to = t[, list(date_time, address = to_address, value)]
balance = rbind(t_to, t_from)
balance = balance[order(date_time)]
balance[, list(Quantity = sum(value)), by=address][order(-Quantity)][1:10]
```

# Question 2: what percentage of the airdrop was claimed so far?

From the blog post (https://uniswap.org/blog/uni/), we know 150 million UNIs are available to be claimed:
"15% of UNI 150,000,000 UNI can immediately be claimed by historical liquidity providers, users, and SOCKS redeemers/holders based on a snapshot ending September 1, 2020, at 12:00 am UTC."


```{r}
t[value==150000000]

t[from_address=="0x090d4613473dee047c3f2706764f49e0821d256e", sum(value)]

```

<div class = "blue">At this time, 114777651 ie **77% of the 150 millions tokens have been claimed**.</div>

How does it look over time? Are we reaching a plateau, with the rest of the claims remaining forgotten?
```{r}
cg = t[from_address=="0x090d4613473dee047c3f2706764f49e0821d256e", sum(value), by=date_time][, list(date_time, total_claimed = cumsum(V1))]
cg[, plot(date_time, total_claimed, type="l", lwd=2, main="Total airdrop tokens claimed")]
legend( "topleft", c("tokens claimed"), text.col=c("black") )
for(i in 0:15) abline(h=i*1e7, col="lightgray")

```

<div class = "blue">
The claims are "plateauing", although there remains about 23% unclaimed.
</div>

This can also be checked on etherscan here:
https://etherscan.io/token/0x1f9840a85d5af5bf1d1762f925bdaddc4201f984?a=0x090d4613473dee047c3f2706764f49e0821d256e#tokenAnalytics


# Question 3: did the receivers of airdrop keep their tokens?

```{r}
claimers_addresses = t[from_address=="0x090d4613473dee047c3f2706764f49e0821d256e", to_address]
length(claimers_addresses)
# we need to remove Binance and Huobi from the list, in case they are claimers
claimers_addresses = claimers_addresses[claimers_addresses != "0x3f5ce5fbfe3e9af3971dd833d26ba9b5c936f0be"] # Binance
claimers_addresses = claimers_addresses[claimers_addresses != "0xe93381fb4c4f14bda253907b18fad305d799241a"] # Huobi
claimers_addresses = claimers_addresses[claimers_addresses != "0x6cc5f688a315f3dc28a7781717a9a798a59fda7b"] # OKEx
length(claimers_addresses)

claimers_trades = t[(to_address %in% claimers_addresses) | (from_address %in% claimers_addresses)]
claimers_trades[(to_address %in% claimers_addresses), value_dir := value]
claimers_trades[(from_address %in% claimers_addresses), value_dir := - value]
claimers_trades[(to_address %in% claimers_addresses) & (from_address %in% claimers_addresses), value_dir := 0]
ctg = claimers_trades[, list(value = sum(value_dir)), by=date_time][order(date_time)]
ctg[, token_count := cumsum(value)]

# Plot
cg[, plot(date_time, total_claimed, type="l", lwd = 2, main="Claimers tokens holdings")]
ctg[, lines(date_time, token_count, type="l", col = "blue", lwd = 2)]
legend( "topleft", c("tokens claimed", "tokens kept"), text.col=c("black", "blue") )
for(i in 0:15) abline(h=i*1e7, col="lightgray")


```

<div class = "blue">
At this time, **claimers of the airdrop have kept (as a group) about 35% of the 115 millions tokens they have claimed**.
They were net sellers of token during the first two or three days; since then they seem to be about neutral (as a group).
</div>

# Question 4: which are the main exchanges?

```{r}
top_from = t[, list(.N, total_tokens_from = sum(value)), by=from_address][order(-total_tokens_from)][N>10][1:5]
top_from$comment = c("Uniswap", "Uniswap: Token Distributor", "Uniswap ETH-UNI pool", "Binance","Huobi 10")
top_from
```

<div class = "blue">
The top markets (exchanges) for UNI tokens so far, are:  the Uniswap ETH-UNI pool, Binance, Huobi.
</div>

```{r}
top_from[3:5, barplot(total_tokens_from, main="Top exchanges of UNI tokens", names.arg=c("Uniswap ETH-UNI pool", "Binance" , "Huobi"), col="darkblue")]
```

# Question 5: who are the main buyers and sellers?

```{r}
top_to = t[, list(.N, total_tokens_to = sum(value)), by=to_address][order(-total_tokens_to)][N>10][1:5]
top_to$comment = c("Uniswap ETH-UNI pool", "Binance", "Huobi 10", "?","OKEx")
top_to

top_addresses_to = t[, .N, by=to_address][order(-N), to_address][1:5]
top_addresses_from = t[, .N, by=from_address][order(-N), from_address][1:5]
top_addresses = unique(c(top_addresses_to, top_addresses_from))
top_addresses
```

<div class = "blue">
The top buyers and sellers are:  the Uniswap ETH-UNI pool, Binance, Huobi, ?, OKEx.
</div>

# Question 6: what is driving the price? (ongoing)

```{r}

# -------------
## Binance trades
# -------------
# We focus on Binance trades, and calculate the cumulative sum of tokens boughts, over time.
binance_trades = t[(to_address=="0x3f5ce5fbfe3e9af3971dd833d26ba9b5c936f0be") | (from_address=="0x3f5ce5fbfe3e9af3971dd833d26ba9b5c936f0be")]
binance_trades[to_address=="0x3f5ce5fbfe3e9af3971dd833d26ba9b5c936f0be", value_dir := value]
binance_trades[from_address=="0x3f5ce5fbfe3e9af3971dd833d26ba9b5c936f0be", value_dir := - value]
bg = binance_trades[, list(value = sum(value_dir)), by=date_time][order(date_time)]
bg[, token_count := cumsum(value)]

# ----------
# Plot
## add extra space to right margin of plot within frame
par(mar=c(5, 4, 4, 6) + 1.5)
bg[, plot(date_time, token_count, col="black", type = "l", xlab="", ylab="", #axes=FALSE, 
          main="Binance UNI tokens held",las=1)]
#axis(2, ylim=c(0,1),col="black",las=1)  ## las=1 makes horizontal labels
mtext("Binance UNI token holdings",side=2,line=4)
abline(h=0)

## Allow a second plot on the same graph
par(new=TRUE)
s[!is.na(r1_mavg), list(r1_mavg = mean(r1_mavg)), by=date_time][, plot(date_time, r1_mavg, xlab="", ylab="",
                        ylim=c(0, 0.03), 
                        type="l",
                        col = "red", axes=FALSE)]
for(i in 0:10) abline(h=i*0.005, col="lightgray")
mtext("UNI vs ETH rate", side=4, col="red", line=4) 
axis(4, col="red",col.axis="red",las=1)



# -------------
## Huobi trades
# -------------
huobi_trades = t[(to_address=="0xe93381fb4c4f14bda253907b18fad305d799241a") | (from_address=="0xe93381fb4c4f14bda253907b18fad305d799241a")]
huobi_trades[to_address=="0xe93381fb4c4f14bda253907b18fad305d799241a", value_dir := value]
huobi_trades[from_address=="0xe93381fb4c4f14bda253907b18fad305d799241a", value_dir := - value]
hg = huobi_trades[, list(value = sum(value_dir)), by=date_time][order(date_time)]
hg[, token_count := cumsum(value)]
```


We will try to show the major token traders activities in one graph together with the token price fluctuation.

```{r, fig.width=14, fig.height=10}
# ----------
# Plot

## add extra space to right margin of plot within frame
par(mar=c(5, 4, 4, 6) + 1.5)
# Binance
bg[, plot(date_time, token_count, col="brown", lwd=2, type = "l", xlab="", ylab="", main="Claimers, Binance, Huobi UNI tokens held",las=1, ylim=c(0,80e6))]

# Huobi
hg[, lines(date_time, token_count, col="blue", lwd=2, type = "l", xlab="", ylab="", las=1)]

# Token claimers (as a group)
ctg[, lines(date_time, token_count, col="darkgreen", lwd=2, type = "l", xlab="", ylab="", las=1)]

mtext("Binance UNI token holdings",side=2,line=4)
abline(h=0)

## Allow a second plot on the same graph, for the token price (in ETH)
par(new=TRUE)
sf = s[!is.na(r1_mavg)]
sf[400:nrow(sf), list(r1_mavg = mean(r1_mavg)), by=date_time][, plot(date_time, r1_mavg, xlab="", ylab="",
                        ylim=c(0, 0.03), 
                        lwd=2, 
                        type="l",
                        col = "red", axes=FALSE)]
for(i in 0:10) abline(h=i*0.005, col="lightgray")
mtext("UNI vs ETH rate", side=4, col="red", line=4) 
axis(4, col="red",col.axis="red",las=1)

legend( "topleft", c("Airdrop claimers tokens", "Binance tokens", "Huobi tokens", "UNI rate vs ETH (right scale)"), 
        text.col=c("darkgreen", "brown", "blue", "red") )

```

The graph is interesting, but the claimers holdings are misleading because they actually were net sellers of tokens, at times.
We should plot instead the net change of holdings of each group, without counting the original claims which were free and could not impact the price in the market.

Let's calculate the net change of holdings for each group:
```{r, fig.width=14, fig.height=14}
# Binance
bg[, netchange100 := frollsum(value, 100)]
# Huobi
hg[, netchange100 := frollsum(value, 100)]

# Claimers
# we need to remove the original claims
ctg = claimers_trades[from_address !="0x090d4613473dee047c3f2706764f49e0821d256e", list(value = sum(value_dir)), by=date_time][order(date_time)]
ctg[, netchange100 := frollsum(value, 100)]


# Plot
## add extra space to right margin of plot within frame
par(mar=c(5, 4, 4, 6) + 1.5)
# Binance
bg[, plot(date_time, netchange100, col="brown", lwd=2, type = "l", xlab="", ylab="", main="Claimers, Binance, Huobi UNI tokens held",las=1, ylim=c(-6e6,6e6))]

# Huobi
hg[, lines(date_time, netchange100, col="blue", lwd=2, type = "l", xlab="", ylab="", las=1)]

# Token claimers (as a group)
ctg[, lines(date_time, netchange100, col="darkgreen", lwd=2, type = "l", xlab="", ylab="", las=1)]

mtext("UNI token holdings",side=2,line=4)
abline(h=0, col="black", lwd=2)


## Allow a second plot on the same graph, for the token price (in ETH)
par(new=TRUE)
sf = s[!is.na(r1_mavg)]
sf[400:nrow(sf), list(r1_mavg = mean(r1_mavg)), by=date_time][, plot(date_time, r1_mavg, xlab="", ylab="",
                        ylim=c(0, 0.03), 
                        lwd=2, 
                        type="l",
                        col = "red", axes=FALSE)]
for(i in 0:10) abline(h=i*0.005, col="lightgray")

mtext("UNI vs ETH rate", side=4, col="red", line=4) 
axis(4, col="red",col.axis="red",las=1)

legend( "topleft", c("Airdrop claimers tokens", "Binance tokens", "Huobi tokens", "UNI rate vs ETH (right scale)"), 
        text.col=c("darkgreen", "brown", "blue", "red") )

```

<div class = "blue">
The analysis of net changes among the top buyers and sellers of tokens suggests the following events (although the analysis is not complete since we only focused on the top 3 groups of buyers and sellers of tokens):

* 17th-18th September 2020: huge volume of claimers of tokens, who unloaded 50% of the tokens to the markets. Binance and Huobi bought most of the tokens available. Price went up.

* 19th September 2020: claims are slowing down. By the end of the day, Binance and Huobi stopped buying, and the claimers were still net sellers. The price started dropping.

* 20th September 2020: volume of trades reduced a lot. Few buyers, and claimers are still net sellers. Price drops further.

* 21st September 2020: price still going down, although it is not clear what drives the price. This would need further investigation.
</div>

### This is the end of the analysis for now.

Thank you.
Fo any question or comment, send me an email at timothee.henry@gmail.com

