---
title: "Uniswap Top pairs using GraphQL and R"
author: "Timothée Henry"
date: "October 8th, 2020"
geometry: margin=1cm
output:
  html_document:
  self_contained: false
lib_dir: libs
---
  
<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>


The objective is to re-create the list of top pairs by liquidity, as displayed here:
https://info.uniswap.org/pairs


```{r}
# ===================
# GraphQL with R
# ===================
# doc:
# https://thegraph.com/explorer/subgraph/uniswap/uniswap-v2
# https://uniswap.org/docs/v2/API/queries/
# ===================
require(ghql)
require(data.table)
library(knitr)
# ==============================
# Most liquid pairs
# ==============================
qry <- Query$new()
qry$query("my_data", 'query {

   __type(name:"Pair") {
      fields {
         name
         description
      }  
   }
}')
con <- GraphqlClient$new('https://api.thegraph.com/subgraphs/name/uniswap/uniswap-v2')
x <- con$exec(qry$queries$my_data)
require(stringr)
#str_split(x, ":null")


qry <- Query$new()
qry$query("my_data", 'query {
 pairs(first: 100, orderBy: reserveUSD, orderDirection: desc) {
          id
          reserveUSD
          volumeUSD
          token0 {
             id
             name
           }
           token1 {
             id
             name
           }
          reserve0
          reserve1
          token0Price
          token1Price
          txCount
          }
          }
          ')

con <- GraphqlClient$new('https://api.thegraph.com/subgraphs/name/uniswap/uniswap-v2')
x <- con$exec(qry$queries$my_data)
t = jsonlite::fromJSON(x)
pairs = data.table(t$data$pairs)
pairs[, reserve0 := round(as.numeric(reserve0))]
pairs[, reserve1 := round(as.numeric(reserve1))]
pairs[, reserveUSD := round(as.numeric(reserveUSD))]
pairs[, volumeUSD := round(as.numeric(volumeUSD))]
pairs[, token0Price := round(as.numeric(token0Price), 2)]
pairs[, token1Price := round(as.numeric(token1Price), 2)]

# re-order the columns
pairs = pairs[, list(id, token0=token0.name, token1=token1.name, txCount, volumeUSD, reserveUSD, token0.id, token1.id, token0Price, token1Price)]

# show top 10 pairs by liquidity (reserveUSD)
kable(pairs[1:10], "pipe")


```

We can verify some numbers on etherscan.io.
The reserves seem correct.
The prices cannot be trusted. Some of the stable coins are priced at 0 instead of 1 USD.
Can we trust the volumeUSD and reserveUSD, then?

To avoid scam tokens, can we get some additional data about the tokens?
For example, I would be interested in:
- number of holders
- number of senders and receivers
- number of days of activity (above a certain $ threshold)


We can do that in BigQuery
```{txt}
SELECT contract_address,
  count(distinct(`from`)) as nb_from,
  count(distinct(`to`)) as nb_to,
  count(distinct( block_timestamp)) as nb_blocks
FROM `blockchain-etl.ethereum_uniswap.UniswapV2Pair_event_Transfer`
group by contract_address
order by nb_to desc
```

and join the data back to our list of pairs:

```{r}
pairs_stats = fread("/Users/henry/Downloads/results-20201008-104503.csv")
setkey(pairs_stats, contract_address)
setkey(pairs, id)


recap = pairs_stats[pairs]
recap[, txCount := as.numeric(txCount)]
recap = recap[order(-txCount)]
recap[, token0Price := NULL] # removing prices which seem wrong
recap[, token1Price := NULL] # removing prices which seem wrong

recap = recap[, list(contract_address, token0, token1, reserve_mUSD = round(reserveUSD/1e6,1), volume_mUSD=round(volumeUSD/1e6,1), txCount, nb_from, nb_to, nb_blocks, token0.id, token1.id)][order(-reserve_mUSD)]
library(knitr)
kable(recap[1:20], "pipe")

```

It is clear now that the reserveUSD cannot be fully trusted. For example the contract 0x6c2b76c8418fc0fcedc26169612ca915e70bd5ea has only 1 transaction, and the transaction is between tokens with no clear value, as per etherscan:
https://etherscan.io/token/0x671d131d72a0fecbd1a5ca9b024a5a150876ba70
https://etherscan.io/address/0x78a685e0762096ed0f98107212e98f8c35a9d1d8

What makes sense at this point is to filter out contracts with too few transactions, and too few participants (distinct to and from addresses).

```{r}

kable(recap[txCount>1000][nb_from>100][nb_to>100][1:50], "pipe")

```



