---
title: "850HPa Temperature Forecast - Persistence"
author: "Timothée Henry"
date: "16/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This analysis uses data and information from:

"Stephan Rasp, Peter D. Dueben, Sebastian Scher, Jonathan A. Weyn, Soukayna Mouatadid, and Nils Thuerey, 2020. WeatherBench: A benchmark dataset for data-driven weather forecasting. arXiv: https://arxiv.org/abs/2002.00469"


```{r}
# ====================
# baseline forecast of temperature 3 days in advance
# compare to https://github.com/pangeo-data/WeatherBench
# ====================

# ------------------------------------------------
# Persistence = forecast previous seen temperature
# ------------------------------------------------
library(ncdf4) # package for netcdf manipulation
library(data.table)

nc_data <- nc_open('/Users/henry/Desktop/perso/climate/WeatherBench/data/temperature_850hPa_2017_5.625deg.nc')
# put temperature variable t in a "d array
t <- ncvar_get(nc_data, "t")
dim(t)

t = as.data.table(t)
setnames(t, c("lon","lat","time","temp"))
```

# Persistence model (same temperature as 3 days before)

```{r}
t[, pred := shift(temp, 72, type="lag"), by=list(lon,lat)]
# rmse
t[!is.na(pred), sqrt(mean((pred - temp)^2))]

# ------------------------------------------------
# RMSE
# ------------------------------------------------
# 4.80 for 2015
# 4.75 for 2016
# 4.80 for 2017
# 4.72 for 2018
# ------------------------------------------------

```

# Analyzing the error of the Persistence model

```{r}
t[!is.na(pred), list(error = sqrt(mean((pred - temp)^2))), by=list(time)][plot(time, error, ylim=c(0,10), main="RMSE by time")]

t[!is.na(pred), list(error = sqrt(mean((pred - temp)^2))), by=list(lat)][plot(lat, error, ylim=c(0,10), main="RMSE by latitude")]

t[!is.na(pred), list(error = sqrt(mean((pred - temp)^2))), by=list(lon)][plot(lon, error, ylim=c(0,10), main="RMSE by longitude")]
```

The RMSE by latitude is clearly the most interesting. Extreme latitudes have highest errors by far.

